---
title: "A computational framework for biomarker discovery in Neuroblastoma through multi-omics data"
subtitle: "Interaction Network Construction"
Author: "Rahma Hussein Hassan Refky"
date: "Jun, 24, 2022"
last modified: "Apr, 20, 2023"
---
#Code for "A computational framework for biomarker discovery in Neuroblastoma through multi-omics data"
#loading libraries:
```{r}
library("rvest")
```
#interactions
```{r}
mirnas <- read.csv("miRNA_list.txt",header = FALSE)
genes <- read.csv("common_gene_list.txt",header = FALSE)


mirnas$V2 <- gsub('mir','miR',mirnas$V1)
minraInGenes <- genes$V1[grepl("MIR",genes$V1)] 
#remove miRNA genes from gene list:
short_genes <- genes$V1[!(genes$V1 %in% minraInGenes)]
```
##transmir:
The evidence level TF-miRNA regulation data for humans was downloaded from the following link: https://www.cuilab.cn/transmir (at the Download tab, H.sapiens (tsv.gz) was downloaded, as it was too large for R to read the file into one dataframe, a simple python code was used to split the file into 13 files with maximum 100000 rows each, the files were named "hsa1.tsv",...,"hsa13.tsv")
here it is:
```{r}
transmir_db <- list()
for (i in 1:13){
  transmir_db[[i]] <- read.delim(file = paste0("E:/transmir_database_evidence_level_hsa/hsa", as.character(i),".tsv", sep=""), header = FALSE, sep = '\t')
}
```
extract interaction data based on our genes and mirnas:
```{r}
transmir_interactions = list()
c=1
for (i in 1:13){
  for (j in 1:dim(transmir_db[[i]])[1]){
    if ((transmir_db[[i]][j,1] %in% short_genes) & (transmir_db[[i]][j,2] %in% mirnas$V1)){
      transmir_interactions[[c]] <- transmir_db[[i]][j,]
      c = c + 1
    }
  }
}


transmir_interactions = as.data.frame(do.call(rbind, transmir_interactions))
colnames(transmir_interactions) <- c("TF name", "miRNA name", "TSS", "Binding site","Action type", "SRAID/PMID", "Evidence", "Tissue", "Species")
```
##tarbase:
I downloaded the latest tarbase datbase version (v8) from here: https://dianalab.e-ce.uth.gr/html/diana/web/index.php?r=tarbasev8%2Fdownloaddataform
```{r}
tarbase_db <- read.table("E:/tarbase/TarBase_v8_download.txt", sep="\t", header = TRUE)
dim(tarbase_db)
tarbase_hsa_db <- tarbase_db[tarbase_db$species == "Homo sapiens",]
dim(tarbase_hsa_db)

colnames(tarbase_hsa_db)
```
extract interactions between our gene/mirna lists from the database:
```{r}
tarbase_interactions <- list()
n <- dim(tarbase_hsa_db)[1]
c <- 1
for (i in 1:n){
  if ((tarbase_hsa_db[i,3] %in% mirnas$V2) & (tarbase_hsa_db[i,2] %in% short_genes)){
    tarbase_interactions[[c]] <- tarbase_hsa_db[i,]
    c <- c + 1
  }
}
tarbase_interactions <- as.data.frame(do.call(rbind, tarbase_interactions))
dim(tarbase_interactions)
```

##combining tarbase and transmir interactions:
```{r}
colnames(transmir_interactions)[c(1,2,9,8)] <- c("source node", "target node", "species", "tissue")
transmir_interactions$`target node` <- gsub("mir", "miR", transmir_interactions$`target node`)
transmir_interactions$species <- "Homo sapiens"
transmir_interactions$database <- "transmir"

tarbase_interactions$database <- "tarbase"
colnames(tarbase_interactions)[c(2,3)] <- c("target node", "source node")

#some renaming to make things easier:

colnames(tarbase_interactions)[c(7,9,10,11)] <- c("cell_line_category","results_type", "method_type", "Action type")
colnames(transmir_interactions)[c(3,4)] <- c("miRNA TSS","TF Binding site")
tarbase_interactions$`Action type`[tarbase_interactions$`Action type` == "DOWN"] <- "Down_regulation"

total_interactions <- plyr::rbind.fill(transmir_interactions, tarbase_interactions)

#replace (,)s with (;)s to make file good:
total_interactions$cell_line_category <- gsub(",",";",total_interactions$cell_line_category)
```
##make node attributes:
```{r}
TFs <- unique(transmir_interactions$`source node`)
GENEs <- unique(tarbase_interactions$`target node`)
MIRNAs <- unique(c(transmir_interactions$`target node`, tarbase_interactions$`source node`))

Nodes <- c(TFs, GENEs[!(GENEs %in% TFs)], MIRNAs)
Molecule_type <- rep(c("TF", "Gene", "miRNA"), c(length(TFs), length(GENEs[!(GENEs %in% TFs)]), length(MIRNAs)))


node_attribs <- data.frame(Nodes, Molecule_type)

interactions <- total_interactions
node_attrib <- node_attribs
```
##retrieving annotations from transmir database url:
```{r}
node_attrib$AssociatedDiseases <- NA
node_attrib$PrognosticCorrelation <- NA
node_attrib$geneId <- NA

interactions_transmir <- interactions[interactions$database == "transmir",]

tfs <- unique(interactions_transmir$`source node`)
for (i in 1:length(tfs)){
  url <- paste0("https://www.cuilab.cn/transmir/tfdetails/H.sapiens&", tfs[i], "/", sep = "")
  temp <- read_html(url)
  tables <- temp %>% html_table()
  node_attrib$geneId[node_attrib$Nodes == tfs[i]] <- tables[[1]][3,2]$X2
  node_attrib$AssociatedDiseases[node_attrib$Nodes == tfs[i]] <- tables[[1]][4,2]$X2
  node_attrib$PrognosticCorrelation[node_attrib$Nodes == tfs[i]] <- tables[[1]][5,2]$X2
}

mirnas <- unique(interactions_transmir$`target node`)
for (i in 1:length(mirnas)){
  url <- paste0("https://www.cuilab.cn/transmir/mirdetails/", gsub("miR", "mir",mirnas[i]), "/", sep = "")
  temp <- read_html(url,config = httr::config(connecttimeout = 60))
  tables <- temp %>% html_table()
  node_attrib$AssociatedDiseases[node_attrib$Nodes == mirnas[i]] <- tables[[1]][4,2]$X2
}

map <- unique(interactions[interactions$database == "tarbase",c("geneId", "target node")])
for (i in 1:dim(node_attrib)[1])
{
  if (node_attrib$Molecule_type[i] == "Gene"){
    node_attrib$geneId[i] <- map[(map$`target node` == node_attrib$Nodes[i]) | (map$`target node` == node_attrib$Nodes[i]), "geneId"]
  }
}
```

##add missing data for transmir interactions:
```{r}
cond1 <- interactions$Evidence == "level 1"
is.na(cond1) <- FALSE
interactions$method[cond1] <- "ChIP-seq"

cond2 <- interactions$Evidence == "level 2"
is.na(cond2) <- FALSE
interactions$method[cond2] <- "ChIP-seq & high-throughput"

#we also have this piece of info: "ChIP is the most direct way to identify binding sites of a single DNA-binding protein or locations of modified histones. " from here: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3591838/#:~:text=ChIP%20is%20the%20most%20direct,or%20locations%20of%20modified%20histones.

#therefore, I'll add the word (direct) to the method type for all chip-seq data:

cond3 <- interactions$Evidence %in% c("level 1", "level 2")
is.na(cond3) <- FALSE
interactions$method_type[cond3] <- "DIRECT"

#unify all na values:
node_attrib[node_attrib == "n/a"] <- NA
interactions[interactions == "n/a"] <- NA
```

##take away level 1 transmir data:
```{r}
cond <- !(interactions$Evidence == "level 1")
cond[is.na(cond)] <- TRUE
interactions_l2 <- interactions[cond,]
nodes <- unique(c(interactions_l2$`source node`, interactions_l2$`target node`))
node_attrib_l2 <- node_attrib[node_attrib$Nodes %in% nodes,]
```
##filtering out interactions found in normal tissues:
```{r}
#filter out interactions in (normal) cell lines:
interactions.filtered = interactions_l2[!(grepl("Normal", interactions_l2$cell_line_category,ignore.case = TRUE)),]

#save filtered interactions:
write.csv(interactions.filtered, file = "FINAL_filtered_Interactions_transmirL2_tarbase_annotated.csv")

nodes_f <- unique(c(interactions.filtered$`source node`, interactions.filtered$`target node`))
node_attrib_f <- node_attrib[node_attrib$Nodes %in% nodes_f,]
write.csv(node_attrib_f, file = "FINAL_filtered_Interactions_transmirL2_tarbase_nodeAttributes_annotated.csv")

#the left out interactions:
interactions.left = interactions_l2[grepl("Normal", interactions_l2$cell_line_category,ignore.case = TRUE),]

#check final numbers of interactions:

#level 2:
dim(interactions_l2)
length(unique(paste0(interactions_l2$`source node`,interactions_l2$`target node`)))
#filtered:
dim(interactions.filtered)
length(unique(paste0(interactions.filtered$`source node`,interactions.filtered$`target node`)))
#normal:
dim(interactions.left)
length(unique(paste0(interactions.left$`source node`,interactions.left$`target node`)))


#check final numbers of nodes:
#level2:
dim(node_attrib_l2)
length(unique(c(interactions_l2$`source node`, interactions_l2$`target node`)))
#filtered:
length(unique(c(interactions.filtered$`source node`, interactions.filtered$`target node`)))
#normal:
length(unique(c(interactions.left$`source node`, interactions.left$`target node`)))

```